----
Machine Learning Course Project: HAR
----

###Synopsis

###Data Processing and Analysis

**Cleaning Data**

Before processing the data it's important to have the data loaded in R.

"classe" is our outcome i.e. variable of interest

Based on the summary of the data set it is possible to conlude:
* dimension of the set: 19622 observations of 160 variables
* time stamp related variables were removed to avoid issues with predictions between test and training sets (different time stamps): raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp
* variables with significant number of NAs were removed from the training set, Where significant was set arbitrary to 80%
* variables with significant number of NAs in the test set were also removed from the training set, Where significant was set arbitrary to 80%
* variables resulting from two above points were cossed checked to choose only the ones common for both test and training sets

```{r}
testData2 <- testData[,-c(3:5)]
trainData2 <- trainData[,-c(3:5)]
trainData2 <- trainData2[,colSums(is.na(trainData2)) < 0.8*nrow(trainData2)]
drop <- testData2[,colSums(is.na(testData2)) >= 0.8*nrow(testData2)]
trainData2 <- trainData2[, !colnames(trainData2) %in% colnames(drop)]

```

**Spending Data**

Create data partition on the training set
```{r}
load(caret)
set.seed(1210)

inTrain <- createDataPartition(trainData2$classe, p = 3/4, list = FALSE)
train <- trainData2[inTrain,]
test <- trainData2[-inTrain,]
myControl <- trainControl(method="cv",number=5,repeats=2,returnResamp='none')
fitModel <- train(classe ~ ., data=train, method="rf", prox=TRUE, trControl=myControl)


```

##vip <- varImp(fitModel$finalModel)
